EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-30 01_intro.tex

%EDIPACK
In this work we present the renewed and improved \NAME{}, version number \textcolor{red}{5.0}: a flexible, high-performance numerical library designed to address these challenges, providing an exact diagonalization (ED) solver for generic Quantum Impurity Problems. 
The library builds on the foundations of
its predecessor~\cite{Amaricci2022}, featuring massively parallel Lanczos-based algorithms while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations for a wide set of impurity models, making it capable of addressing a large set of problems, from spin-orbit coupling effects in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact
dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

@@

%MC this last point is intrinsically DMFT (or embedding) and it has no connection with impurity physics
This new version also includes support for systems with non-equivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites. This capability makes it 
possible to explore a wide range of inhomogeneous systems, including
oxide interfaces or disordered systems, where
translational symmetry is broken.

@@

%\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of \NAME \textcolor{red}{5.0}, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging impurity problems. 
The rest of this work has the following structure. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration
and installation procedures. In the following  \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in full detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface which is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) a basic Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and 
all the discussed extensions with a series of illustrative examples.

>>>>

%EDIPACK
In this work we present the renewed and improved \NAME{}: a flexible, high-performance numerical library designed to address these challenges, providing an exact diagonalization (ED) solver for generic Quantum Impurity Problems. 
The library builds on the foundations of
its predecessor~\cite{Amaricci2022}, featuring massively parallel Lanczos-based algorithms while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations for a wide set of impurity models, making it capable of addressing a large set of problems, from spin-orbit coupling effects in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact
dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

@@

%MC this last point is intrinsically DMFT (or embedding) and it has no connection with impurity physics
This new version also includes support for systems with non-equivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites. This capability makes it 
possible to explore a wide range of inhomogeneous systems, including
oxide interfaces or disordered systems, where
translational symmetry is broken.
%Repetition

@@

%\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of \NAME \textcolor{red}{2.0 i get that we want to say this is the updated edipack, but now it is version 5.0.5... i believe this could create confusion}, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging impurity problems. 
The rest of this work has the following structure. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration
and installation procedures. In the following  \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in full detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface which is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) a basic Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and 
all the discussed extensions with a series of illustrative examples.
